Scaling Automated Theory Exploration

### 2015-08-06 AI4FM/article.tex
Fixed underscores

% Article frontmatter
% ===================
\title{Scaling Automated Theory Exploration} % Title of the article
%\short{} % Short title of the article (optional)
\author{Chris Warburton\sponsor{}} % Authors and references to addresses
\institute{ % Institutes with labels
\email{cmwarburton@dundee.ac.uk}, \url{http://www.computing.dundee.ac.uk/about/staff/124}\\
School of Computing\\
University of Dundee}
\abstract{
We investigate the \textbf{theory exploration} (TE)
paradigm for computer-assisted Mathematics and identify limitations and
improvements for current approaches. Unlike the theorem-proving paradigm,
which requires user-provided conjectures, TE performs an open-ended
search for theorems satisfying given criteria. We see promise in TE for
identifying new abstractions and connections in libraries of software
and proofs, but realising this potential requires more scalable
algorithms than presently used.} % Abstract of the article
%\keywords{} % Keywords for the article

@@

In existing systems, this criterion is intimately connected to the
search algorithm; this coupling improves efficiency, even for brute-force
search. As an example, \textsc{IsaCoSy} \cite{johansson2009isacosy} discovers
equations, which are defined as ``interesting'' if they cannot be simplified
using previously discovered equations. The intuition for such criteria is to
avoid special cases of known theorems, such as \(0 + 0 = 0\), \(0 + 1 = 1\),
etc. when we already know \(\forall x. 0 + x = x\). To work effectively, this
requires general forms to be found \emph{before} special cases.

@@

\section{Tackling Complexity}

\section{Existing Work}

@@

\section{Going Forward}

@@

@@

% Acknowledgements for colleagues, referees, ...
% ==============================================
\begin{acknowledge}
I am grateful for those who have helped formulate these ideas through
conversation, especially the HipSpec team at Chalmers University (Moa
Johansson, Koen Claessen, Nick Smallbone, Dan Rosén and Irene Lobo Valbuena)
and my supervisor Katya Komendantskaya. I also wish to thank the implementors
of the systems we are building on, including HipSpec, QuickSpec and
QuickCheck on the theory exploration side, as well as GHC, Cabal and Nix
on the infrastructure side.
\end{acknowledge}

>>>>

% Article frontmatter
% ===================
\title{Scaling Automated Theory Exploration} % Title of the article
%\short{} % Short title of the article (optional)
\author{Chris Warburton\sponsor{}} % Authors and references to addresses
\institute{ % Institutes with labels
\email{cmwarburton@dundee.ac.uk}\\
School of Computing\\
University of Dundee}
\abstract{
We investigate the \textbf{theory exploration} (TE)
paradigm for computer-assisted Mathematics and identify limitations and
improvements for current approaches. Unlike the theorem-proving paradigm,
which requires user-provided conjectures, TE performs an open-ended
search for theorems satisfying given criteria. We see promise in TE for
identifying new abstractions and connections in libraries of software
and proofs, but realising this potential requires more scalable
algorithms than presently used.} % Abstract of the article
%\keywords{} % Keywords for the article

@@

In existing systems, this criterion is intimately connected to the
search algorithm; this coupling improves efficiency, even for brute-force
search. As an example, \textsc{IsaCoSy} \cite{johansson2009isacosy} discovers
equations, which are defined as ``interesting'' if they cannot be simplified
using previously discovered equations. The intuition for such criteria is to
avoid special cases of known theorems, such as $0 + 0 = 0$, $0 + 1 = 1$,
etc. when we already know $\forall x. 0 + x = x$. To work effectively, this
requires general forms to be found \emph{before} special cases.

@@

In the following, we give an overview of the state of the art in automated
theory exploration; identify potential improvements; and present our initial
attempts to implement some of these ideas.

\section{Theory Exploration in Haskell}

@@

\section{Identifying Improvements}

@@

\section{The \textsc{ML4HS} Framework}

@@

% Acknowledgements for colleagues, referees, ...
% ==============================================
\begin{acknowledge}
I am grateful for those who have helped formulate these ideas through
conversation, especially the HipSpec team at Chalmers University (Moa
Johansson, Koen Claessen, Nick Smallbone, Dan Ros{\'e}n and Irene Lobo Valbuena)
and my supervisor Katya Komendantskaya. I also wish to thank the implementors
of the systems we are building on, including HipSpec, QuickSpec and
QuickCheck on the theory exploration side, as well as GHC, Cabal and Nix
on the infrastructure side.
\end{acknowledge}
